A Meta-analysis of Longevity Estimates of Mosquito Vectors of Disease: Supplementary Material

### 2018-11-23 supplementary_materials.tex
Read up to `Here we estimate our model with one of six different survival functions '

The data collection method resulted in 568 separate dissection series, across 72 published articles (see Section \ref{sec:appendix_dissectionStudyList} for a list of the studies included in our final dataset). The published datasets cover the period from 1960-2015 with comparable numbers of studies across each decade (Fig. \ref{fig:dissection_timeSeries}). The statistical approach applied to the data relies on the assumption that there is a constant rate of recruitment into the adult population (see Section \ref{sec:dissection_dissectionStats}). If populations fluctuate strongly from month-to-month this assumption will likely be violated. To try to mitigate against such a risk, we aggregate the data across all dates to obtain a single series for each identifier, resulting in 201 such series. To obtain a reasonable level of accuracy on estimates we remove all those individual (aggregated) series where there are fewer than 100 mosquitoes in total in the series. Finally we remove data for any species where there was only one series resulting in 131 series across four genera (\textit{Anopheles}, \textit{Culex}, \textit{Aedes}, and \textit{Mansonia}) and 25 species (Table \ref{tab:dissection_speciesNumbers}). These studies are distributed across a wide range of geographies (Fig. \ref{fig:dissection_map}). The raw data for the dissection analysis are shown in Figures \ref{fig:dissection_exampleData1}, \ref{fig:dissection_exampleData2} and \ref{fig:dissection_exampleData3}.

@@

\subsection{Statistical analysis of dissection data}\label{sec:dissection_dissectionStats}
We suppose that the number of individuals recruited into the adult female population is constant over time, meaning that the age-structure of the population is stable. Whilst environmental heterogeneity will naturally lead to variation in adult recruitment over time, we hope that by aggregating data across studies undertaken over a range of dates we may reduce the impact of this effect on our results. Suppose that the probability an individual female mosquito survives until age $a$ is given by the survival function $S(a)$, then the number of individuals in the population surviving to this age is given by,
%
\begin{align}\label{eq:dissection_structured}
A(a) = A(0) S(a),
\end{align}
%
where $A(0)$ is the number of female mosquitoes recruited into the population per unit time. Consider one individual experiment where we randomly sample from a wild population that is structured as per eq. (\ref{eq:dissection_structured}). In this case the number of individuals sampled at each age is binomially-distributed,
%
\begin{align}
X(a) \sim \mathcal{B}(A(a),p(a)),
\end{align} 
%
where $p(a)$ is the probability of recapturing a single mosquito of age $a$. In what follows we typically assume that the probability of capturing a given mosquito is independent of their age, so that $p(a)=p=const$. Since in general $A(a)$ is large and $p(a)$ is small, we can approximate the above using a Poisson distribution,
%
\begin{align}
X(a) \sim \text{Poisson}(A(a)p).
\end{align}
%
However we believe that the assumption of \textit{independent} captures of individual mosquitoes, which underlies the binomial and Poisson models, is likely suspect for the same reasons as for the MRR analysis (see Section \ref{sec:mrr_statistical}). As before we choose to specify a negative binomial sampling distribution that allows for non-independent captures,
%
\begin{align}
X(a) \sim \text{NB}(A(a)p,\kappa),
\end{align} 
%
where we use the parameterisation such that the mean is given by $A(a)p$, and the over-dispersion parameter, $\kappa$, where as $\kappa\rightarrow \infty$ the above sampling distribution approaches a Poisson.

@@

Here we estimate our model with one of six different survival functions (the same as for the MRR case; Table \ref{tab:mrr_survivalDescription}), each of which makes different assumptions regarding how the force of mortality is affected by mosquito age. To estimate lifespan at the species, genus and overall groupings we then use a hierarchical Bayesian model of the same mathematical form as for the analysis of MRR experiments (see Section \ref{sec:MRR_hierarchical}). However the priors for the analysis of dissection data were modified so that they represented a mean prior lifespan of three gonotrophic cycles, although allowed considerable variation in mean lifespan (Fig. \ref{fig:dissection_lifespanPriorsParity}). The priors used for the over-dispersion parameter ($\kappa$) for the negative binomial likelihood were the same as for the MRR analysis.

>>>>

The data collection method resulted in 568 separate dissection series, across 72 published articles (see Section \ref{sec:appendix_dissectionStudyList} for a list of the studies included in our final dataset). The published datasets cover the period from 1960-2015 with comparable numbers of studies across each decade (Fig. \ref{fig:dissection_timeSeries}). The statistical approach applied to the data relies on the assumption that there is a constant rate of recruitment into the adult population (see Section \ref{sec:dissection_dissectionStats}). If populations fluctuate strongly from month-to-month this assumption will likely be violated. To try to mitigate against such a risk, we aggregate the data across all dates to obtain a single series for each identifier, resulting in 201 such series. To obtain a reasonable level of accuracy on estimates we remove all those individual (aggregated) series where there are fewer than 100 mosquitoes in total in the series. Finally we remove data for any species where there was only one series resulting in 131 series across four genera (\textit{Anopheles}, \textit{Culex}, \textit{Aedes}, and \textit{Mansonia}) and 25 species (Table S\ref{tab:dissection_speciesNumbers}). These studies are distributed across a wide range of geographies (Fig. \ref{fig:dissection_map}). The raw data for the dissection analysis are shown in Figures \ref{fig:dissection_exampleData1}, \ref{fig:dissection_exampleData2} and \ref{fig:dissection_exampleData3}.

@@

\subsection{Statistical analysis of dissection data}\label{sec:dissection_dissectionStats}
We suppose that the number of individuals recruited into the adult female population is constant over time, meaning that the age-structure of the population is stable. Whilst environmental heterogeneity will naturally lead to variation in adult recruitment over time, we hope that by aggregating data across studies undertaken over a range of dates we may reduce the impact of this effect on our results. Suppose that the probability an individual female mosquito survives until age $a$ is given by the survival function $S(a)$, then the number of individuals in the population surviving to this age is given by,
%
\begin{align}\label{eq:dissection_structured}
A(a) = A(0) S(a),
\end{align}
%
where $A(0)$ is the number of female mosquitoes recruited into the population per unit time. Consider one individual experiment where we randomly sample from a wild population that is structured as per eq. (\ref{eq:dissection_structured}). In this case the number of individuals sampled at each age is binomially-distributed,
%
\begin{align}
X(a) \sim \mathcal{B}(A(a),p(a)),
\end{align} 
%
where $p(a)$ is the probability of recapturing a single mosquito of age $a$. In what follows we assume that the probability of capturing a given mosquito is independent of their age, so that $p(a)=p=const$. Since in general $A(a)$ is large and $p(a)$ is small, we can approximate the above using a Poisson distribution,
%
\begin{align}
X(a) \sim \text{Poisson}(A(a)p).
\end{align}
%
However we believe that the assumption of \textit{independent} captures of individual mosquitoes, which underlies the binomial and Poisson models, is likely suspect for the same reasons as for the MRR analysis (see Section \ref{sec:mrr_statistical}). As before we choose to specify a negative binomial sampling distribution that allows for non-independent captures,
%
\begin{align}
X(a) \sim \text{NB}(A(a)p,\kappa),
\end{align} 
%
where we use the parameterisation such that the mean is given by $A(a)p$, and the over-dispersion parameter, $\kappa$, where as $\kappa\rightarrow \infty$ the above sampling distribution approaches a Poisson.

@@

Here we estimate our model with one of six different survival functions (the same as for the MRR case; Table S\ref{tab:mrr_survivalDescription}), each of which makes different assumptions regarding how the force of mortality is affected by mosquito age. To estimate lifespan at the species, genus and overall groupings we then use a hierarchical Bayesian model of the same mathematical form as for the analysis of MRR experiments (see Section \ref{sec:MRR_hierarchical}). However the priors for the analysis of dissection data were modified so that they represented a mean prior lifespan of three gonotrophic cycles, although allowed considerable variation in mean lifespan (Fig. \ref{fig:dissection_lifespanPriorsParity}). The priors used for the over-dispersion parameter ($\kappa$) for the negative binomial likelihood were the same as for the MRR analysis.
